\section*{Course Overview}

Welcome! We're very excited about your interest in this course.

Gravwell exists to provide data analytics to organizations at a
predictable and reasonable cost that enables any business unit to gain
insights from data. Founded in 2017, Gravwell was created to address
some core deficiencies in the market as experienced by our founders.
From the very first line of code they knew people needed a tool
that was much more efficient on computing resources, could handle binary
data natively in order to shrink disk usage and support advanced machine
learning, and didn't charge for every ounce of data like some kind of
analytics butcher shop.

One incredibly important aspect of data analytics is the ability to ask
ad-hoc questions and to perform rapid data exploration. This requires an
``ingest first and ask questions later'' approach to data ingestion,
indexing, and querying. Additionally, data trends over time are
incredibly important for identifying cybersecurity incidents, user
behavior analytics, and to extract business KPIs and insights. Gravwell
needed a structure-on-read time-series database, but a suitable option
didn't really exist. In order to accomplish our mission, Gravwell
engineers had to start from scratch to create a full-stack analytics
platform built with Go. The success of this bold undertaking is what
allows Gravwell to scale so incredibly well and to support such a wide
variety of use cases.

This training is a deep dive into the Gravwell platform. We will cover
the platform architecture, data ingestion strategies, user management
and permissions, automation capabilities, cluster management and
troubleshooting, and of course, searching. Lots of searching. This
document serves as a reference guide and detailed walkthrough for this
course.

We really appreciate your interest in Gravwell. We strive to provide
amazing customer service and that extends into this training material.
If you have any questions, comments, or suggestions, please do send them
to us at feedback@gravwell.io. Your opinion is extremely important
to us.

Thanks again, and enjoy the training!

Sincerely,


The Gravwell Team



